Air cargo load and route planning in pickup and delivery operations

### 2024-02-04 latex/ESWA.tex
Time complexity

\tikzset{
	mybox/.style={rectangle,
		draw,
		fill= myblue,
		rounded corners,
		minimum width=2cm,
		inner sep=5pt,
		align=left,
		minimum height=1cm
	},
	myarrow/.style={draw=black,
		fill=white,
		minimum width=0.6cm,
		single arrow
	},
	longarrow/.style={draw=none,
		shading=axis,
		left color=white,
		right color=myblue,
		minimum width=0.6cm,
		single arrow,
		anchor=east
	}
}

@@

\begin{abstract}
In the aerial pickup and delivery of goods in a distribution network, transport aviation faces risks of load imbalance due to the urgency required for loading, immediate take-off, and mission accomplishment. Transport planners deal with trip itineraries, prioritisation of items, building up pallets, and balanced loading, but there are no commercially available systems that can integrally assist in all these requirements. This enables other risks, such as improper delivery, excessive fuel burn, and possible safety issues due to cargo imbalance, as well as a longer than necessary turn-around time. This {\it NP-hard}\/ problem, named {\it Air Cargo Load Planning with Routing, Pickup, and Delivery Problem (ACLP+RPDP)}, is mathematically modelled using standardised pallets in fixed positions. We developed a strategy to solve this problem, considering historical transport data from some Brazilian hub networks, and performed several experiments with a commercial solver, four known meta-heuristics, and a new heuristic designed specifically for this problem. By using a portable computer, our strategy quickly found practical solutions to a wide range of real problems in much less than operationally acceptable time.
\end{abstract}

@@

A few years ago, \citep{BrandtStefan2019} defined the {\it Air Cargo Load Planning Problem} (ACLPP) as four sub-problems: {\it Aircraft Configuration Problem} (ACP), {\it Build-up Scheduling Problem} (BSP), {\it Air Cargo Palletization Problem} (APP), and {\it Weight and Balance Problem} (WBP). Several aspects were considered: item characteristics to be transported (dimensions, scores, dangerousness, etc.); types and quantities of {\it unit load devices} (ULDs); when these pallets are assembled; how items are allocated to pallets; in which positions these pallets are to be placed; how total cargo weight is balanced; etc.

{\color{blue}

Recently, \citep{zhao2023} developed integer programming solution that address ACLPP and WBP separately, impacting payload optimization and aircraft centre of gravity (CG).
The authors argue that, with improvements in computer processing power, joint combinatorial optimization of ACPP and WBP is now feasible. They propose three integer linear programming models: a bi-objective optimization model (BOM), a combinatorial optimization model (COM), and an improved combinatorial optimization model (IOM). The models aim to determine the maximum loading capacity and lowest CG deviation from a specified target CG.

The study uses Gurobi to solve four scenarios with various conditional metrics for three models. Results show that the BOM has the fastest solution speed, but the CG deviation is the largest, making the results unacceptable in some cases. The COM has the longest solution time, making it difficult to tolerate in practice. The IOM offers the best optimization solution, despite taking a little longer to solve computationally than the BOM.

Using a Boeing 777F, four scenarios and 25 cases were tested, finding that the BOM's variable feasible domain is limited, the optimization result is insufficient, and the CG deviation is sometimes too large. The COM takes too long to solve, making it difficult to meet the needs of the actual scene. The IOM accelerates the solving process compared to the COM, but some scenarios still require significant computation time.

{\color{violet}

Inefficient air transport plans can lead to unnecessary costs, extra routes, longer distances, and incorrect destinations. Unbalanced cargo increases fuel consumption due to altered aircraft pitch angles.

A study by \cite[p. 22]{gwh2007} highlights the risk of weight- and balance-related accidents on cargo flights compared to passenger flights. So, balancing cargo is crucial for safe aerial transportation, as an improperly positioned centre of gravity can result in dangerous take-off and landing conditions and stall recovery issues.

Another important issue is managing fuel consumption. To get an idea of the influence of CG displacement on fuel consumption, \cite[p. 140]{MongeauBes2003} report that {\it "a displacement of the CG of less than 75 cm in a long-range aircraft yields, over a 10,000 km flight, a saving of 4,000 kg of fuel"}. This reinforces the importance of balancing cargo.

Despite technological advancements, many airlines still rely on manual aircraft loading and balancing, using simple computer-aided tools for CG computation and constraint testing. Loadmasters often create load plans manually, which is time-consuming and can lead to flight delays.

As to the right delivery, an organisation's purchasing department may acquire supplies in the main capitals of the country and abroad to meet the company's needs, and it wants these materials to be delivered to the requesting organisations within the stipulated deadlines.

A rational decision-making process is crucial to avoid creating inefficient or unsafe transport plans, considering the high costs of fuel, maintenance, operation, outsourcing expenses, potential operational impairments, and safety risks due to unbalanced cargo.

Solving the ACLPP+RPDP is vital for optimising strategic scores of what must be transported, saving time and effort in loading, ensuring safety and balance, ensuring correct pickups and deliveries, and finding the best routes considering fuel use due to potential cargo imbalance.

@@

@@

The vast majority of operational research applied to air cargo is focused on challenges related to WBP, that is, the distribution of items on pallets to ensure load balancing. We can mention: \citep{LarsenMikkelsen1979}; \citep{Brosh1981}; \citep{Kevin1992}; \citep{Heidelberg1998}; \citep{fok2004optimizing}; \citep{KaluznyBohdanL2009Oalb}; \citep{Verstichel2011}; \citep{Limbourg2012}; \citep{RoesenerBarnes2016}; \citep{YangLiuGao2018}; \citep{zhao2021}; \citep{MiguelOptimalAP}.

Other authors have addressed pallet assembly (APP) on aircraft, possibly also considering load balancing (WBP): \citep{MongeauBes2003}; \citep{Chan2006}; \citep{RoesenerHall2014}; \citep{Vancroonemburg2014}; \citep{Paquay2016}; \citep{Paquay2018}; \citep{wong2020}; \citep{eugene2021}; \citep{zhao2023}.

In all these works, there is a great diversity of scenarios and solutions: some consider items in two dimensions, and others in three dimensions; some used integer programming, and others developed specific heuristics.

\citep{LurkinSchyns2015} is the only work that simultaneously addresses an air cargo (WBP) and a flight itinerary (PDP) sub-problem. The authors demonstrated that this problem is {\it NP-hard}. Although it is innovative, strong simplifications were imposed by these authors: in relation to loading, APP was ignored; regarding routing, it is assumed that a predefined tour plan is restricted to only two legs. Referring directly to this work, \citep{BrandtStefan2019} comment: {\it However, not even these sub-problems are acceptably solved for real-world problem sizes, or models omit some practically relevant constraints}.

@@

\begin{table}[htp]
	\centering
	\caption{Air cargo transport: literature, sub-problems and features}  \label{tab:sa}
	\footnotesize
	\begin{tabular}{r|cccc|ccccc}
		\toprule
		Work & {\bf APP}  & {\bf WBP}  &  {\bf PDP}   &{\bf TSP}   & {\bf 2D}  & {\bf 3D}  & {\bf Heu}  & {\bf Int}  & {\bf Lin} \\
		\midrule
		\citep{LarsenMikkelsen1979}  & $.$        & $\bigstar$ & $.$          & $.$        & $.$       & $.$       & $\bigstar$ & $.$        &  $.$ \\
		\citep{Brosh1981}  & $.$ & $\bigstar$  & $.$   & $.$ & $.$ & $.$   & $.$  & $.$  &  $\bigstar$ \\
		\citep{Kevin1992}  & $.$ & $\bigstar$  & $.$ & $.$ & $.$ &$.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{Heidelberg1998}  & $.$ & $\bigstar$  & $.$ & $.$ & $\bigstar$ &$.$   & $\bigstar$  & $.$  &  $.$ \\
		\citep{MongeauBes2003}    & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{fok2004optimizing} & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\	
		\citep{Chan2006}  & $\bigstar$ & $.$    & $.$ & $.$ & $.$ & $\bigstar$  & $\bigstar$  & $.$  &  $.$ \\
		\citep{KaluznyBohdanL2009Oalb}  & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$ &$.$  & $.$  & $\bigstar$  &  $.$ \\
		\citep{Verstichel2011}   & $.$ & $\bigstar$    & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\	
		\citep{Limbourg2012} & $.$ & $\bigstar$  & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{RoesenerHall2014}  & $\bigstar$ & $\bigstar$  & $.$  & $.$ & $.$ & $\bigstar$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{Vancroonemburg2014}  & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{LurkinSchyns2015} & $.$ & $\bigstar$  & $\bigstar$ & $.$  & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\citep{RoesenerBarnes2016}  & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $\bigstar$  & $.$  &  $.$ \\
		\citep{Paquay2016,Paquay2018}  & $\bigstar$ & $\bigstar$ & $.$ & $.$ & $.$ & $\bigstar$ & $\bigstar$  & $\bigstar$ & $.$ \\
		\citep{YangLiuGao2018} & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$  & $.$ & $\bigstar$ & $.$  & $.$ \\
		\citep{wong2020} & $\bigstar$  & $\bigstar$  & $.$  & $.$   & $.$  & $.$ & $.$ & $\bigstar$  & $.$  \\
		\citep{eugene2021} & $\bigstar$ & $\bigstar$ & $.$  & $.$   & $.$ & $.$ & $.$ & $\bigstar$  & $.$  \\
		\citep{zhao2021} & $.$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$ \\
		\citep{zhao2023} & $\bigstar$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$  \\
		\citep{MiguelOptimalAP} & $.$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$  \\
		{\bf This article}   & $\bigstar$ & $\bigstar$  & $\bigstar$& $\bigstar$ & $.$ & $.$ & $\bigstar$ & $\bigstar$   &  $.$  \\
		\bottomrule 
	\end{tabular}
	\normalsize 
\end{table}

@@

{\color{blue}

As for the size of the problem dealt with in this work, addressing the scale of the Travelling Salesperson Problem (TSP) in practical, real-world air transport contexts indeed reveals a fascinating and complex challenge. In the realm of academic research, TSP problems often explore theoretical models that can scale to hundreds or even larger numbers of cities.

In practical terms, especially within the air transport sector, the size and complexity of a TSP can vary significantly depending on several factors:

\textbf{Fleet and Route Planning}: For airline operations, the problem isn't just about finding the shortest path through a set of points; it also involves scheduling, aircraft assignment, crew scheduling, maintenance planning, and delivering the right cargo to the right client. This last operation task complexity transforms a standard TSP into a more intricate problem in our study, the Pickup and Delivery Problem (PDP).

\textbf{Operational Constraints}: Factors like weather, air traffic control restrictions, and the availability of take-off and landing slots make real-world TSP scenarios in air transport even more challenging. Although we do not deal with these constraints, they add layers of complexity that significantly depart from the theoretical TSP model.

\textbf{Dynamic Nature}: Unlike the static nature of classical TSP, real-world problems in air transport are highly dynamic. Routes need to be optimized continuously as conditions change, requiring algorithms that can adapt in real-time or near-real-time.

In practical air transport scenarios, the size of a TSP can be less about the sheer number of nodes and more about the complexity and dynamic nature of the problem, which in our case (a single aircraft in a single day) is rarely expected to be more than seven nodes. Despite this, at the end of this work, we to submit the best method found to a more challenging 15-node problem as a validation test.
}

@@

\vspace{-2mm}

@@

We consider real-world scenarios, where Table \ref{tab:larger} shows the aircraft parameters. $p_i$\/ are pallets, $1\leq i \leq 18$, whose weight and volume limits are $W_i$\/ and $V_i$, respectively. $D_i^{long}$\/ and $D_i^{lat}$\/ are, respectively, the longitudinal and lateral distances of each pallet centroids to the aircraft centre of gravity (CG) along both axes. These distances will be used in the calculation of the torque, referring to the items allocated on each pallet. In this aircraft, as the ramp has an inclination of $25^{\circ}$, we made the necessary corrections in $D_i^{long}$, $W_i$\/ and $V_i$\/ of the corresponding pallets ($p_1$, $p_2$, $p_3$, and $p_4$).

@@

Once the assumptions and the mathematical modelling are presented, it is possible to see that ACLP+RPDP is {\it NP-hard}. In a similar way to \citep{LurkinSchyns2015}, consider the simple case where $K=1$\/ (one leg), $m=2$\/ (two pallets around the aircraft CG), $2n$\/ sufficiently light items with same scores in node 0, and no items in node 1. Under these conditions, through polynomial reductions for the {\it Set-Partition Problem}, it is possible to demonstrate that the decision problem associated with ACLP+RPDP is {\it NP-complete}. For more details, see \citep[p.~6]{LurkinSchyns2015}.

@@

In practical cases, we know that a common aircraft has $m=18$\/ pallets, flight itineraries have $K \leq 6$\/ nodes plus the base, and each node has hundreds of items to be shipped. We also know that missions with fewer nodes are more frequent than longer ones. Under these circumstances, we can adopt some important strategies:

\vspace{-2mm}

	\item We consider that the number of destinations is smaller than the number of pallets, that is, $K < m$. We also avoid the trivial case where $K=1$. With this premise, we can preset the destinations of the pallets at each shipping node, reserving for each destination a number of pallets proportional to the volume available. We could have used another criterion, but it was observed in the experiments that the volume is more constrictive in airlift.

{\color{violet}
	\item As the number of nodes is small, we have the possibility to check all possible tours, selecting the one that provides the best value for the objective function. For this reason, a maximum time per solution ($tmax$) is established. In the tests carried out in this work, these times were in the range [240s, 1200s, 2400s, 3600s].

	\item At each node, we will use a MIP solver or heuristics to check which method offers the best result within the time limit for each node.

	\item The time limit to solve a node will be adjusted to be proportional to the sum of volumes of the candidate items of the node in relation to the sum of volumes of the entire tour.
}

@@

Our complete strategy is summarized in Algorithm \ref{alg:main}.

@@

	\begin{algorithmic}[1]
			\State {\bf ACLP+RPDP} \,\, {\it in:}~$scenario, surplus, tmax, tsp, ntours$ \, \, {\it out:}~$answer$

@@

		{\color{violet}
			\If {$tsp$ = True}

				\State $tmax \gets tmax/ntours$ \Comment{max. tour time to solve the number of tours from the TSP heuristic}

			\Else

				\State $tmax \gets tmax/K!$

			\EndIf
		}
			\For {each $method$}  \label{main:method}

				\For {each $\pi \in S_K$} \label{main:loop1}

					\State $f_{\pi} \gets SolveTour(\pi, L, M, C, N, method, tmax)$ \label{main:fpi}

@@

				\State $answer[scenario,surplus,method] \gets \max f_{\pi}$ \label{main:a2}
			\EndFor
	\end{algorithmic}

@@

In this algorithm, we use {\color{blue}six} values for $scenario$, according to Tables \ref{tab:costs} and \ref{tab:scenarios}, which define the number $K$ of destinations, the set $L$\/ of nodes, the costs $C$\/, and the shortest tours $\pi_{TSP1}$\/ and $\pi_{TSP2}$.

{\color{blue}
Scenario 6 will be used only for the best method validation.
}

\begin{figure}[htp]

	\centering
	\fbox{
	\begin{tikzpicture}

		%% Make the boxes
		\node [mybox] (first)                        { \small {\bf Select:}\\ 
			\small 1. scenario\\ 
			\small 2. surplus\\ 
			\small 3. time limit};

		\node [mybox, right=0.7cm of first] (second) { \small {\bf Config:}\\ 
			\small 4. parameters\\ \small 5. cost matrix\\ 
			\small 6. pallets\\ 
			\small 7. items};

		\node [mybox, right=0.7cm of second] (third) { \small {\bf Define tour}\\  
			\small {\bf schema:}\\ 
			\small 8. All K! tours\\ \ \ \  
			\small or\\ 
			\small 8. Solve a TSP\\ 
			\small  \ \ \ (set of tours))};

>>>>

\tikzset{
	mybox/.style={rectangle,
		draw,
		fill= myblue,
		rounded corners,
		minimum width=2cm,
		inner sep=5pt,
		align=left,
		minimum height=1cm
	},
	myarrow/.style={draw=black,
		fill=white,
		minimum width=0.6cm,
		single arrow
	},
	downarrow/.style={draw=black,
	fill=white,
	minimum width=0.6cm,
	single arrow
	},
	longarrow/.style={draw=none,
		shading=axis,
		left color=white,
		right color=myblue,
		minimum width=0.6cm,
		single arrow,
		anchor=east
	}
}

@@

\begin{abstract}
In the aerial pickup and delivery of goods in a distribution network, transport aviation faces risks of load imbalance due to the urgency required for loading, immediate take-off, and mission accomplishment. Transport planners deal with trip itineraries, prioritisation of items, building up pallets, and balanced loading, but there are no commercially available systems that can integrally assist in all these requirements. This enables other risks, such as improper delivery, excessive fuel burn, and possible safety issues due to cargo imbalance, as well as a longer than necessary turn-around time. This {\it NP-hard}\/ problem, named {\it Air Cargo Load Planning with Routing, Pickup, and Delivery Problem (ACLP+RPDP)}, is mathematically modelled using standardised pallets in fixed positions. We developed a strategy to solve this problem, considering historical transport data from some Brazilian hub networks, and performed several experiments with a commercial solver, {\color{blue}five} known meta-heuristics, and a new heuristic designed specifically for this problem. By using a portable computer, our strategy quickly found practical solutions to a wide range of real problems in much less than operationally acceptable time.
\end{abstract}

@@

A few years ago, \cite{BrandtStefan2019} defined the {\it Air Cargo Load Planning Problem} (ACLPP) as four sub-problems: {\it Aircraft Configuration Problem} (ACP), {\it Build-up Scheduling Problem} (BSP), {\it Air Cargo Palletization Problem} (APP), and {\it Weight and Balance Problem} (WBP). Several aspects were considered: item characteristics to be transported (dimensions, scores, dangerousness, etc.); types and quantities of {\it unit load devices} (ULDs); when these pallets are assembled; how items are allocated to pallets; in which positions these pallets are to be placed; how total cargo weight is balanced; etc.

@@

Inefficient air transport plans can lead to unnecessary costs, extra routes, longer distances, and incorrect destinations. Unbalanced cargo increases fuel consumption due to altered aircraft pitch angles, increasing the risk of weight- and balance-related accidents. Balancing cargo is crucial for safe aerial transportation, as an improperly positioned centre of gravity (CG) can result in dangerous take-off and landing conditions and stall recovery issues. Despite technological advancements, many airlines still rely on manual aircraft loading and balancing, which can lead to flight delays. A rational decision-making process is essential to avoid creating inefficient or unsafe transport plans, considering high costs of fuel, maintenance, operation, outsourcing expenses, potential operational impairments, and safety risks due to unbalanced cargo. Solving this problem is vital for optimizing strategic scores, saving time and effort in loading, ensuring safety and balance, correct pickups and deliveries, and finding the best routes considering fuel use due to potential cargo imbalance.

@@

The vast majority of operational research applied to air cargo is focused on challenges related to WBP, that is, the distribution of items on pallets to ensure load balancing. We can mention: \cite{LarsenMikkelsen1979}; \cite{Brosh1981}; \cite{Kevin1992}; \cite{Heidelberg1998}; \cite{fok2004optimizing}; \cite{KaluznyBohdanL2009Oalb}; \cite{Verstichel2011}; \cite{Limbourg2012}; \cite{RoesenerBarnes2016}; \cite{YangLiuGao2018}; \cite{zhao2021}; \cite{MiguelOptimalAP}.

Other authors have addressed pallet assembly (APP) on aircraft, possibly also considering load balancing (WBP): \cite{MongeauBes2003}; \cite{Chan2006}; \cite{RoesenerHall2014}; \cite{Vancroonemburg2014}; \cite{Paquay2016}; \cite{Paquay2018}; \cite{wong2020}; \cite{eugene2021}; \cite{zhao2023}.

In all these works, there is a great diversity of scenarios and solutions: some consider items in two dimensions, and others in three dimensions; some used integer programming, and others developed specific heuristics. Among the most recent, we can highlight:
\begin{itemize}

	\item \cite{RoesenerBarnes2016} proposed a heuristic to solve the {\it Dynamic Airlift Loading Problem} (DALP). Given a set of palletized cargo items that require transport between two nodes in a given time frame, the objective of this problem is to select an efficient subset of aircraft, partition the pallets into aircraft loads, and assign them to allowable positions in those aircraft.

	\item \cite{Paquay2016} presented a mathematical model to optimise the loading of heterogeneous 3D boxes on pallets with a truncated parallelepiped format. Its objective is to maximise the volume used in containers, considering load balancing constraints, the presence of fragile items, and the possibility of rotating these boxes. \cite{Paquay2018} developed some heuristics to solve this problem.

	\item \cite{YangLiuGao2018} modelled the air transport problem as a 2D packing problem and presented a heuristic for its optimisation in several aircraft, considering load balancing to minimise fuel consumption.

	\item \cite{wong2020} developed a mathematical model and a tool based on mixed integer programming for optimising cargo in aircraft with different pallet configurations. Balance constraints and the presence of dangerous items were considered. \cite{eugene2021} integrated this tool into a digital simulation model with a visualisation and validation system based on sensors that alert about load deviations.

	\item \cite{zhao2021} proposed a model for WBP based on Mixed Integer Programming (MIP). Instead of focusing on the CG deviation, the authors consider the original CG envelope of the aircraft, with a linearization method for its non-linear constraints.

	\item \cite{zhao2023} proposed three integer linear programming models: a bi-objective optimisation model (BOM), a combinatorial optimisation model (COM), and an improved combinatorial optimisation model (IOM). Its objectives were the maximum possible loading and the smallest CG deviation. Four scenarios with multiple metrics are resolved for the {\it Boeing 777F}. The results showed that the BOM has the fastest solution speed, but the CG deviation is the worst; in several cases, the CG deviation is unacceptable. COM has the longest solution time, which is difficult to tolerate in practice.

\cite{LurkinSchyns2015} is the only work that simultaneously addresses an air cargo (WBP) and a flight itinerary (PDP) sub-problem. The authors demonstrated that this problem is {\it NP-hard}. Although it is innovative, strong simplifications were imposed by these authors: in relation to loading, APP was ignored; regarding routing, it is assumed that a predefined tour plan is restricted to only two legs. Referring directly to this work, \cite{BrandtStefan2019} comment: {\it However, not even these sub-problems are acceptably solved for real-world problem sizes, or models omit some practically relevant constraints}.

@@

\begin{table}[htp]
	\centering
	\caption{Air cargo transport: literature, sub-problems and features}  \label{tab:sa}
	\footnotesize
	\begin{tabular}{r|cccc|ccccc}
		\toprule
		Work & {\bf APP}  & {\bf WBP}  &  {\bf PDP}   &{\bf TSP}   & {\bf 2D}  & {\bf 3D}  & {\bf Heu}  & {\bf Int}  & {\bf Lin} \\
		\midrule
		\cite{LarsenMikkelsen1979}  & $.$        & $\bigstar$ & $.$          & $.$        & $.$       & $.$       & $\bigstar$ & $.$        &  $.$ \\
		\cite{Brosh1981}  & $.$ & $\bigstar$  & $.$   & $.$ & $.$ & $.$   & $.$  & $.$  &  $\bigstar$ \\
		\cite{Kevin1992}  & $.$ & $\bigstar$  & $.$ & $.$ & $.$ &$.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{Heidelberg1998}  & $.$ & $\bigstar$  & $.$ & $.$ & $\bigstar$ &$.$   & $\bigstar$  & $.$  &  $.$ \\
		\cite{MongeauBes2003}    & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{fok2004optimizing} & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\	
		\cite{Chan2006}  & $\bigstar$ & $.$    & $.$ & $.$ & $.$ & $\bigstar$  & $\bigstar$  & $.$  &  $.$ \\
		\cite{KaluznyBohdanL2009Oalb}  & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$ &$.$  & $.$  & $\bigstar$  &  $.$ \\
		\cite{Verstichel2011}   & $.$ & $\bigstar$    & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\	
		\cite{Limbourg2012} & $.$ & $\bigstar$  & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{RoesenerHall2014}  & $\bigstar$ & $\bigstar$  & $.$  & $.$ & $.$ & $\bigstar$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{Vancroonemburg2014}  & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{LurkinSchyns2015} & $.$ & $\bigstar$  & $\bigstar$ & $.$  & $.$ & $.$   & $.$  & $\bigstar$  &  $.$ \\
		\cite{RoesenerBarnes2016}  & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $\bigstar$  & $.$  &  $.$ \\
		\cite{Paquay2016,Paquay2018}  & $\bigstar$ & $\bigstar$ & $.$ & $.$ & $.$ & $\bigstar$ & $\bigstar$  & $\bigstar$ & $.$ \\
		\cite{YangLiuGao2018} & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$  & $.$ & $\bigstar$ & $.$  & $.$ \\
		\cite{wong2020} & $\bigstar$  & $\bigstar$  & $.$  & $.$   & $.$  & $.$ & $.$ & $\bigstar$  & $.$  \\
		\cite{eugene2021} & $\bigstar$ & $\bigstar$ & $.$  & $.$   & $.$ & $.$ & $.$ & $\bigstar$  & $.$  \\
		\cite{zhao2021} & $.$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$ \\
		\cite{zhao2023} & $\bigstar$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$  \\
		\cite{MiguelOptimalAP} & $.$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$  \\
		{\bf This article}   & $\bigstar$ & $\bigstar$  & $\bigstar$& $\bigstar$ & $.$ & $.$ & $\bigstar$ & $\bigstar$   &  $.$  \\
		\bottomrule 
	\end{tabular}
	\normalsize 
\end{table}

@@

@@

@@

We consider real-world scenarios, where Table \ref{tab:larger} shows the aircraft parameters. $p_i$\/ are pallets, $1\leq i \leq 18$, whose weight and volume limits are $W_i$\/ and $V_i$, respectively. $D_i^{long}$\/ and $D_i^{lat}$\/ are, respectively, the longitudinal and lateral distances of each pallet centroids to the aircraft CG along both axes. These distances will be used in the calculation of the torque, referring to the items allocated on each pallet. In this aircraft, as the ramp has an inclination of $25^{\circ}$, we made the necessary corrections in $D_i^{long}$, $W_i$\/ and $V_i$\/ of the corresponding pallets ($p_1$, $p_2$, $p_3$, and $p_4$).

@@

Once the assumptions and the mathematical modelling are presented, it is possible to see that ACLP+RPDP is {\it NP-hard}. In a similar way to \cite{LurkinSchyns2015}, consider the simple case where $K=1$\/ (one leg), $m=2$\/ (two pallets around the aircraft CG), $2n$\/ sufficiently light items with same scores in node 0, and no items in node 1. Under these conditions, through polynomial reductions for the {\it Set-Partition Problem}, it is possible to demonstrate that the decision problem associated with ACLP+RPDP is {\it NP-complete}. For more details, see \cite[p.~6]{LurkinSchyns2015}.

@@

\begin{figure}[htp]
	\centering
	\fbox{ \small
		\begin{tikzpicture}
			\node [mybox] (one) { \small {\bf Input data:}\\
				\small \ \ \ {Aircraft parameters}\\
				\small \ \ \ {Airports and distances}\\
				\small \ \ \ {$scenario$ \ \ ($K$, $L$\/ and $C$)}\\
				\small \ \ \ {Items available for shipment at each airport}\\						
				\small \ \ \ {$surplus = \sum_{j=1}^{n_k} v_j$/$\sum_{i=1}^{m} V_i$ }\\		
				\small \ \ \ {$tmax$ \ \ (overall runtime limit)}\\
				\small {{\bf Requirement:} $1 < K < m$}
			};

			\node [mybox, below=0.7cm of one] (two) { \small {\bf Number of tested tours:}\\
				\small \ \ \ {{\bf if} $K \leq 6$}\\
				\small \ \ \ \ \ \ {$ntours = 2$ \ \ (optimal TSP solutions)}\\
				\small \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\it or}\\				
				\small \ \ \ \ \ \ {$ntours = K!$ \ \ (all possible tours)}\\
				\small \ \ \ {\bf else}\\
				\small \ \ \ \ \ \ {$ntours = 100$ \ \ (best tours obtained with a TSP heuristic)}
			};

			\node [mybox, below=0.7cm of two] (three) { \small {\bf On each tour of $ntours$:}\\
				\small \ \ \ {Calculate a node-by-node solution (divide runtime according to shipment volumes):}\\
				\small \ \ \ \ \ \ \ {- Reallocate packed contents with torque optimization}\\
				\small \ \ \ \ \ \ \ {- Preset pallet destinations according to shipment volumes}\\
				\small \ \ \ \ \ \ \ {- Find a node solution (MIP, Shims, NMO, ACO, GRASP, TS, GA) }\\
				\small \ \ \ \ \ \ \ {- Accumulate values obtained from score, torque and cost}\\
				\small \ \ \ \ \ \ \ {- Go to the next node on the current tour}
			};

In practical cases, we know that a common aircraft has $m=18$\/ pallets, flight itineraries have $K \leq 6$\/ nodes plus the base, and each node has hundreds of items to be shipped. We also know that missions with fewer nodes are more frequent than longer ones. Under these circumstances, we can adopt some important strategies summarized in Figure \ref{fig:process}:

	\item We consider that the number of destinations is smaller than the number of pallets ($K < m$), and we avoid the trivial case where $K=1$. With this premise, we can preset the destinations of the pallets at each shipping node, reserving a number of pallets proportional to the volume available for each destination. We could have used another criterion, but it was observed in the experiments that the volume is more constrictive in airlift.

	\item An important parameter is the number $ntours$\/ of tours tested. In practical cases where $K \leq 6$, we have the possibility to check all possible tours ($ntours = K!$). In this situation, as $K$\/ is small, we can also specially analyse the two optimal solutions of the corresponding TSP ($ntours = 2$). Finally, in cases where $K>6$, we will use a heuristic to select $100$\/ tours of low length ($ntours=100$), and search among them for the one that provides the best value for the objective function.

	\item To compare the performance of each strategy, an overall runtime limit $tmax$\/ is established and divided by $ntours$\/ tours. In turn, the runtime limit for each tour will be distributed among its nodes in proportion to the volume available for boarding.

	\item At each node of a tour, the packed contents that remain on board are reallocated on pallets in order to minimize torque on the aircraft. This calculation is done quickly using a MIP solver. Then, the destinations of the pallets are previously defined in proportion to the shipment volume. Finally, considering the runtime limit of each node, we will use a MIP solver and five well-known meta-heuristics to find the best allocation of shipping items: {\it Ant Colony Optimization} (ACO), {\it Noising Method Optimization } (NMO), {\it Tabu Search} (TS), {\it Greedy Randomized Adaptive Search Procedure} (GRASP), and {\it Genetic Algorithm} (GA).

	\item We will generate benchmarks using the $surplus$\/ parameter, which is a value in $\{1.2, 1.5, 2.0\}$. It corresponds, at each node $k$, to the ratio between the sum of the volumes of the items and the load capacity of the pallets ($surplus = \sum_{j=1}^{n_k} v_j$/$\sum_{i=1}^{m} V_i$). This parameter allows us to verify the different behaviour of each method, according to $scenario$\/ and the quantity of items available for shipment.

	\item We will do tests by varying the number $K$\/ of destinations, the set $L$\/ of nodes, and the costs $C$. Each group of values tested is called $scenario$, according to Tables \ref{tab:costs} and \ref{tab:scenarios}, where $1<K \leq 6$. After finding the method with the best performance in node-by-node solution, we will test it in solving cases with $K > 6$.

@@

Algorithm \ref{alg:main} is the main program of this solution process, which parameters are $scenario$, $surplus$, $tmax$, and $ntours$. The input data is obtained from Tables \ref{tab:costs}, \ref{tab:larger} and \ref{tab:scenarios} (lines \ref{main:M}-\ref{main:KLC}).

@@

	\begin{algorithmic}[1]
			\State {\bf ACLP+RPDP} \,\, {\it in:}~$scenario, surplus, tmax, ntours$ \, \, {\it out:}~$answer$

@@

			\For {each $method$}  \label{main:method}

				\For {$x \gets 1$ to $ntours$} \label{main:loop1}

					\State $f(S_K[x]) \gets SolveTour(S_K[x], L, M, C, N, method, tmax/ntours)$ \label{main:fpi}

@@

				\State $answer[scenario,surplus,method] \gets \max f(S_K[x])$ \label{main:a2}
			\EndFor
	\end{algorithmic}

@@

$surplus$ is passed to $ItemsGeneration$\/ (line \ref{main:items}), responsible for creating the items to be shipped, which will be presented in the next section (Algorithm \ref{alg:itemsgen}). $tmax$\/ is the runtime limit, which will be distributed among the tours (line \ref{main:fpi}). $method$\/ corresponds to a MIP solver or a heuristic to the node-by-node solution $SolveTour$, which will be presented in subsection \ref{methods}. The best results obtained by testing all tours are stored in $answer$\/ (line \ref{main:a2}), which is the output of this algorithm.

{\color{red}$\pi_k$ are tours} {\color{blue}$\pi$ are tours}, that is, permutations with the $K$\/ nodes. When $ntours=2$, $\pi_1$\/ and $\pi_2$\/ are the optimal solutions of the corresponding TSP. When $ntours=K!$, $\pi_k$\/ is the $k^{th}$ permutation of $S_K$. Finally, when $ntours=100$, $\pi_1, \ldots, \pi_{100}$\/ are the best TSP solutions provided by a heuristic.
